### 2015-04-15 docs/ms.Rmd
Updated text to accepted version, excluding the results section.

# Reproducible research in the study of biological colouration

@@

Data archiving: Github repo (http://dx.doi.org/10.5281/zenodo.13245)

@@

The study of colour in nature has generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010]. The ongoing interest in biological colouration has in some part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [e.g. @endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003]. While these developments are positive for the field, the increasingly complex analyses being run on ever greater amounts of data heighten the need for comprehensive methods reporting and diligent data management [@alsheikh_public_2011; @nekrutenko_next_2012].

Replication and transparency lie at the heart of the scientific enterprise. Beyond simply allowing independent verification of results, reproducible research ensures greater comparability between studies and provides a foundation for testing new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007]. A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly available and archived, and (iii) analysis code is publicly available, such that the chain of modification of raw data is documented and preserved. While completely reproducible research [e.g. @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable. Nevertheless, even partial reproducibility through the relatively simple practices of complete methods reporting and public data archiving is of tremendous value.

Our aim was to explore the state of reproducibility in the study of biological colouration, and to outline simple ways in which it may be improved. We begin by outlining common methods for studying biological colouration, with an emphasis on the way in which their subtle complexity can dramatically shape results. We then present guidelines for the comprehensive reporting of methods, and analyse a subset of the literature against these criteria to assess the state of methods reporting. We also quantify the availability of publicly archived data and code, and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.

@@

Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and uncovering best practices in measuring colour has been a great challenge. The application of spectrometry (also called spectroradiometry, or spectrophotometry) to questions of biological colouration revolutionized the field [@dyck_determination_1966]. Now relatively inexpensive and convenient, spectrometry allows for the rapid capture of quantitative data, and has been widely adopted as standard across the field [@andersson_quantifying_2006]. Digital photography is also increasingly being used to quantify biological colouration [@stevens_using_2007], as high-resolution cameras are inexpensive and allow for the simultaneous, repeatable sampling of multiple colour patches [@mckay_use_2013].

Subtle variation in the application of these common methods can significantly alter the resulting data. When using spectrometry, for example, the angle from which spectra are collected [@santos_strong_2007], the calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen will all directly affect the recorded spectra [@andersson_quantifying_2006; @dalrymple_roses_2014; @meadows_quantifying_2011; @schaefer_birds_2008]. Equally, the main challenge for gathering colour data from photographs is ensuring standardization against nonlinear responses to light intensity, biases towards particular wavebands, and inconsistent lighting [@stevens_using_2007]. Also, most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system. Thus hardware modifications for ultraviolet and infra-red imaging are often necessary [e.g. @stevens_colour_2013].

Expansion in the availability of objective methods for the measurement of colour has been matched by advances in theory and analysis. In particular, the development of visual models has enabled researchers to move beyond purely quantitative comparisons of reflectance spectra and adopt more biologically relevant perspectives when defining and testing hypotheses [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998]. Visual models typically attempt to describe the reception and early-stage processing of colour information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g. @endler_comparing_2005; @vorobyev_receptor_1998]. Although relatively easy to implement, visual models are built on multiple assumptions about the way in which stimuli are processed that can dramatically shape the results of a given analysis [@lind_avian_2009; @pike_preserving_2012]. For example, Vorobyev et al.’s [-@vorobyev_receptor_1998; -@vorobyev_colour_2001] widely used receptor-noise model requires researchers to specify the form of receptor quantum catches (i.e. raw versus transformed), the use of any colour constancy mechanism (e.g. von Kries), as well as the photoreceptor signal-to-noise ratios, their relative densities, and the type of noise predominating (i.e. neural and/or receptor noise). These data are only available for a small number of species, such as humans and honeybees, and so assumptions and extrapolations are necessary when models are applied to other species. Unless the biological rationale underlying every parameter choice is clearly understood, justified, and reported, such analyses may become ineffective, especially when comparing results between studies [@pike_preserving_2012].

The comprehensive reporting of methods is a simple and crucial step in ensuring research is reproducible. Accordingly, we developed a list of essential information about the capture (Table 1) and analysis (Table 2) of colour data that should ideally be reported. Analytical techniques are diverse, mathematically complex, and are being developed rapidly [@kelber_spectral_2010]. Rapid progress means that the need for a deep understanding of common methods can quickly outstrip the working knowledge of the average researcher. As a consequence, the subtle complexity of many analytical techniques can be overlooked by empiricists, leading to critical methodological information not being reported. Our guidelines for reporting the details of colour analyses (Table 2) thus cover two broad, common methods: colourimetric (or 'spectral') analyses, and visual modelling. These tables present the framework for our subsequent analysis of methods reporting in the literature.

@@

To assess the current state of reproducibility in the field we searched papers from 2013 in 22 leading journals: _American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_. On each of the journals' homepages, we used the Boolean phrase 'colour\*' or 'color\*' or 'spectra\*' to search the title or abstract. To review the 216 papers returned from our search, journals were haphazardly divided up between the authors. We excluded review papers, methodological papers, papers quantifying spatial rather than chromatic properties of a colour patch, and studies taking micro-spectrometric measurements of retinal absorbance. The final set of `r nrow(coldat)` papers included only those that used either a spectrometer or camera to quantify colouration. In order to reduce the risk of observer bias in our assessment, each paper included in the final set was read and reassessed by three authors. Any discrepancies between assessor scores were discussed and resolved prior to analysis. We then tabulated what details of the methods were reported against our suggested criteria (Tables 1 and 2), and recorded whether data (in either a 'raw' or 'processed' form) and/or any code were publicly available. The data along with analysis script has been stored as a github repo (http://dx.doi.org/10.5281/zenodo.13245). We have kept the papers used in our dataset anonymous as our aim was to explore the general question of reproducibility in the field and not single out particular studies.

@@

The types of analysis varied across studies. Overall, `r analysisType[1]` papers generated colourimetric variables [i.e. 'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used both colourimetrics and visual models, and `r analysisType[3]` used other forms of analysis (e.g. purely statistical analyses, such as principle component analysis on reflectance spectra). Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models alone, and `r analysisTypeSpec[3]` used other analyses. Studies using colourimetric measures specifically defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 1)`% of the studies.

@@

Of the `r nrow(coldat)` studies analysed, `r round((data["raw"]/sum(data))*100, digits = 1)`% publicly provided the raw underlying data, `r round((data["proc"]/sum(data))*100, digits = 1)`% provided data in a pre-processed form, and `r round((data["none"]/sum(data))*100, digits = 1)`% of studies did not provide any publicly accessible data. The paucity of data being made available post-publication in studies of biological colouration is in line with other fields [@drew_lost_2013; @vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014]. There is evidence, however, that this trend is shifting as funding agencies and publishers increasingly mandate the release of data [@whitlock_data_2010]. The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods. This is particularly relevant in the study of colour traits, where new methods for analysing colour data are frequently developed [e.g. @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014]. More generally, the provision of open data may also foster collaborations as researchers draw upon existing work [@whitlock_data_2011], and considerably reduce the difficulty of meta-analyses and large scale comparative work [e.g. @burd_distinctive_2014; @maia_key_2013]. Increased data sharing has been shown to positively correlate with citations [@piwowar_sharing_2007], and given the increasing rate of retractions there is also impetus for researchers to maintain credibility through transparency [@steen_retractions_2010; @van_trouble_2011; @wicherts_willingness_2011]. We thus encourage researchers to publicly archive data in as raw a form as possible (e.g. individual reflectance spectra), to maximise its utility. Public data repositories such as figshare and dryad allow the permanent archival of data as citable research objects, and the preparation of data for publication is often straightforward [@whitlock_data_2011]. The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data and are subsequently credited, though we acknowledge that tensions over data-sharing exist [discussed in @roche_troubleshooting_2014].

None of the included studies linked to any form of code, which is not surprising given the popularity of graphical-user-interface (GUI) based statistical and colour-analytical software [e.g. AVICOL; @gomez_avicol_2006]. Here we simply note that there are several advantages to moving away from a GUI-based to a programming workflow, in spite of the initial time investment required. Programming languages such as Python [@van_python_1995] and R [@R_program], for example, are free, open-source, and are host to communities of developers that continuously build and implement analytical tools.  Indeed, tools specifically for the analysis of colour data already exist [e.g. the package 'pavo' for R; @maia_pavo_2013]. Programming languages are flexible by nature, which allows individual researchers to rapidly explore new methods as they are published, and methods developers to implement their analyses to increase the reach and impact of their work. A significant advantage of programmatic analyses is that analysis scripts, when properly curated, represent a complete documented history of a study's methods. The chain of modification of raw data may thus be preserved, both for researchers revisiting their own work in the future and for other scientists looking to build directly on the results of previous studies.

@@

Our review of the recent literature highlights some significant impediments to reproducibility in the study of biological colouration. The simplest step towards reproducible research is through the comprehensive reporting of methods, yet crucial methodological information is often unreported (Fig. 1).  In particular, the light source - sample - probe geometry was rarely reported, yet should be standard reporting for all studies using spectrometry. Visual models were regularly used to analyse spectral data. However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of noise used in receptor-noise models (Fig. 1). Photoreceptor densities should also be reported where possible even if using standard avian densities, as all of these parameters have important consequences for downstream analysis and, hence, the reproducibility of colour trait quantification [e.g. @kemp_ornamental_2008; @santos_comparison_2007]. We also recommend the presentation of spectra where possible, as these figures allow for the rapid assessment of the nature and quality of the data independent of downstream processing. We found that only `r round(propSpec_fin["prop_refl_fig"], digits = 1)`% of studies using spectral reflectance data had such a figure. Although not considered in our analysis, we also recommend that authors explicitly outline the biological justification underlying their choices for data capture and analysis. The permanent, public storage of data also plays a pivotal role in reproducibility and transparency, yet public data storage in studies of colour traits is not yet the norm. We encourage researchers to consider the substantial benefits of publicly archiving their data [@piwowar_sharing_2007; @vision_open_2010; @whitlock_data_2010].

Our objective in this paper was to highlight some important gaps that can influence the reproducibility of colour research, and to offer guidelines on simple ways in which studies may be leveraged to their full potential. We do not wish to deprecate studies that happen to lack some methodological details. We hope that our discussion of methods reporting and data archiving will contribute to the growing culture of reproducibility in science, and that researchers will consider these issues when deciding on the best practices in their work. It is only through critical appraisal that we can identify areas for improvement, and constructively contribute to the growth of our field. Such assessments are particularly important in an age where accountability, transparency and reproducibility are at the forefront of discussions of the state of scientific research [@drew_lost_2013; @peng_reproducible_2011; @wicherts_willingness_2011]. We believe that part of the issue at hand is simply that researchers seldom have clear guidelines to follow when undertaking their research, and may overlook important aspects of methods and analysis reporting. We hope our critical evaluation of recent literature has highlighted a need for more careful descriptions of how colour is measured and analysed. We are optimistic that this may help to establish uniform reporting guidelines for publication and increase transparency and reproducibility, which will continue the advance of an exciting era in colour research.

>>>>

# Reproducible research in the study of biological coloration

@@

Data archiving: Github repo (?)

@@

The study of colour in nature has generated insights into fundamental evolutionary and ecological processes, and research into colour traits is a rapidly growing field [@kelber_spectral_2010]. The ongoing interest in biological coloration has in part been driven by the increased availability of key technologies, including spectrometry and photography, and concurrent advances in methods for analysing colour data, such as visual models [e.g., @endler_comparing_2005; @stevens_using_2007; @kelber_animal_2003]. While these developments are positive for the field, the increasingly complex analyses being run on ever-greater amounts of data heighten the need for comprehensive methods reporting and diligent data management [@alsheikh_public_2011; @nekrutenko_next_2012].

Replication and transparency lie at the heart of the scientific enterprise. Beyond simply allowing independent verification of results, reproducible research ensures greater comparability between studies and provides a foundation for testing new ideas and methods [@whitlock_data_2011; @van_trouble_2011; @piwowar_sharing_2007]. A study may be considered truly reproducible when it satisfies three broad criteria: (i) methods are reported completely, (ii) data are publicly available and archived, and (iii) the chain of modification of raw data is documented and preserved. While completely reproducible research [e.g., @fitzjohn_much_2014] is a laudable goal, the considerable demands it imposes on researchers means that it will often, in practice, be unattainable. Nevertheless, even partial reproducibility through the relatively simple practices of complete methods reporting and public data archiving is of tremendous value.

Our aim was to explore the state of reproducibility in the study of biological coloration, and to outline simple ways in which it may be improved. We first outline common methods for studying biological coloration and present guidelines for comprehensive methods reporting. We then explore how well some of these important criteria are reported in the literature. We also quantify the availability of publicly archived data and code and suggest some useful tools for increasing the reproducibility of colour trait research more broadly.

@@

Generations of biologists have endeavoured to explain the mechanisms and functions of animal and plant colouration [@endler_comparing_2005; @thayer_concealing_1909; @poulton_colours_1890; @wallace_natural_1891], and uncovering best practices in measuring colour has been a great challenge. The direct measurement of reflectance and/or transmittance through spectrometry revolutionized the study of biological coloration [@dyck_determination_1966], and has been widely adopted as the standard [@andersson_quantifying_2006]. Digital photography is also increasingly being used to quantify colour [@stevens_using_2007], as high-resolution cameras are inexpensive and allow for the simultaneous, rapid sampling of multiple colour patches [@mckay_use_2013].

Expansion in the availability of objective methods for the measurement of colour has been matched by advances in theory and analysis. In particular, the development of visual models has enabled researchers to move beyond quantitative comparisons of reflectance spectra and adopt potentially more biologically relevant perspectives when defining and testing hypotheses [@chittka_colour_1992; @endler_comparing_2005; @vorobyev_receptor_1998]. Visual models typically attempt to describe the reception and early-stage processing of chromatic and achromatic information as a function of an object’s reflectance, the ambient illumination, and a receiver’s sensory system [e.g., @endler_comparing_2005; @vorobyev_receptor_1998]. Although relatively easy to implement, visual models are built on multiple assumptions about the way in which stimuli are processed that can dramatically shape the results of a given analysis [e.g., @lind_avian_2009; @pike_preserving_2012].

## Guidelines for methods reporting

Given that methodological variation may shape results in significant and unpredictable ways (Tables 1 & 2, and references therein), the comprehensive reporting of methods is a simple and crucial step in ensuring research is reproducible. Accordingly, we developed a list of information about the capture (Table 1) and analysis (Table 2) of colour data that should ideally be reported. With regards to the measurement of colour, we focus on the two most frequently used methods; photography and spectrometry. Analytical techniques are diverse, mathematically complex, and are being developed rapidly [@thery_insect_2010; @kelber_spectral_2010]. Such progress means that the need for a deep understanding of common methods can quickly outstrip the working knowledge of the average researcher. As a consequence, the subtle complexity of many analytical techniques can be overlooked by empiricists, leading to critical methodological information not being reported. Our guidelines for reporting the details of colour analyses (Table 2) thus cover two broad, common methods: colourimetric (or 'spectral') analyses, and visual modelling.

While we wish to emphasize that these details are essential to ensuring full reproducibility of data capture and analysis, we recognize that space restrictions in the main text of manuscripts may preclude the incorporation of all these details. In such cases, we suggest that details be included in meta-data or in a supplementary file so that the necessary information is available to researchers. It is also the case that there is variability in the degree to which each parameter may affect results, and so we have aimed to provide a brief, qualitative outline of the potential effects that variation in each parameter may have on the data (Table 1 & 2). These tables are not intended as a guide to the selection of methods, however, for which we refer readers to excellent recent reviews as well as the original publications [@stevens_using_2007; @kelber_animal_2003; @montgomerie_analyzing_2006; @kemp_integrative_2015, and references in Tables 1 & 2].

@@

To determine the current state of reproducibility in the field we assessed a sample of the literature against a set of our criteria (Table 1 and 2), which we expected should be commonly reported based on our background reading. We searched papers from 2013 in 22 leading journals: _American Journal of Botany_, _The American Naturalist_, _Animal Behaviour_, _Behavioral Ecology_, _Behavioral Ecology and Sociobiology_, _Biological Journal of the Linnean Society_, _Biology Letters_, _Current Zoology_, _Ecology_, _Ecology and Evolution_, _Ecology Letters_, _Ethology_, _Evolution_, _Functional Ecology_, _Journal of Ecology_, _The Journal of Evolutionary Biology_, _The Journal of Experimental Biology_, _Naturwissenschaften_, _New Phytologist_, _Oikos_, _PLoS ONE_, _Proceedings of the Royal Society B: Biological Sciences_. On each of the journals' homepages, we used the Boolean phrase 'colour*' or 'color*' or 'spectra*' to search the title and/or abstract. Journals were haphazardly divided up between the authors to review the 216 papers returned from our search. On first pass, we excluded review papers, methodological papers, papers quantifying spatial (i.e., pattern) rather than chromatic properties of a colour patch, and studies taking micro-spectrometric measurements of retinal absorbance. The final set of 60 papers included only those that used either a spectrometer or camera to quantify coloration. In order to reduce the risk of observer bias in our assessment, each paper included in the final set was read and reassessed by two further authors. Any discrepancies between assessment scores were discussed by the three authors that had read the article and resolved prior to analysis. We also recorded whether data (in either a 'raw' or 'processed' form) and/or any code were publicly available. The data along with our analysis script has been stored as a github repository. We have kept the papers used in our dataset anonymous as our aim was to explore the general question of reproducibility in the field.

@@

The types of analysis varied across studies. Overall, `r analysisType[1]` papers generated colourimetric variables [i.e. 'spectral' measures of hue, saturation, and/or brightness; @montgomerie_analyzing_2006] whereas `r analysisType[4]` used visual models, `r analysisType[2]` used both colourimetrics and visual models, and `r analysisType[3]` used other forms of analysis (e.g., purely statistical analyses, such as principle component analysis on reflectance spectra). Of the studies quantifying colour using spectral reflectance, `r analysisTypeSpec[1]` used colourimetric measures, `r analysisTypeSpec[2]` used multiple analyses, `r analysisTypeSpec[4]` used visual models alone, and `r analysisTypeSpec[3]` used other analyses. Studies using colourimetric measures specifically defined their measure of brightness, hue and/or chroma in `r round(col_def, digits = 1)`% of the studies.

@@

Of the `r nrow(coldat)` studies analysed, `r round((data["raw"]/sum(data))*100, digits = 1)`% publicly provided the raw underlying data, `r round((data["proc"]/sum(data))*100, digits = 1)`% provided data in a pre-processed form, and `r round((data["none"]/sum(data))*100, digits = 1)`% of studies did not provide any publicly accessible data. The paucity of data being made available post-publication in studies of biological coloration is in line with other fields [@drew_lost_2013; @vines_mandated_2013; @wolkovich_advances_2012], including the broader field of animal behaviour [@caetano_forgotten_2014]. There is evidence, however, that this trend is shifting as funding agencies and publishers increasingly mandate the release of data [@whitlock_data_2010].

The clearest benefit of open access to data is that it allows researchers to build upon previous work for the testing and refinement of new ideas and methods. As discussed above, it is also essential that complete methodological details be provided, either in the manuscript or associated meta-data, to ensure data can be used to their full potential. This is particularly relevant in the study of colour, where new analysis methods are frequently developed [e.g., @allen_analyzing_2013; @endler_framework_2012; @stoddard_pattern_2014]. More generally, the provision of open data is likely to foster collaborations as researchers draw upon existing work [@piwowar_sharing_2007; @whitlock_data_2011], and considerably reduce the difficulty of meta-analyses and large scale comparative work [e.g., @burd_distinctive_2014; @maia_key_2013]. We thus encourage researchers in our field to publicly archive raw data (e.g., individual reflectance spectra), to maximise its utility. The use of modest data embargoes and appropriate licenses can help to ensure that original authors are able to make the best use of data and are subsequently credited, though we acknowledge that tensions over data-sharing exist [discussed in @roche_troubleshooting_2014].

None of the included studies linked to any form of code. This is not surprising given the popularity of graphical-user-interface (GUI) based statistical and colour-analytical software [e.g., AVICOL; @gomez_avicol_2006]. When such software is freely available, and is combined with the comprehensive reporting of methods, its use represents a valuable approach to conducting reproducible research.  We also note that there are advantages in adopting a programming workflow, in spite of the initial time investment required. Programming languages such as Python [@van_python_1995] and R [@R_program], for example, are free, open-source, and are host to communities of developers that continuously build and implement analytical tools.  Indeed, tools specifically for the analysis of colour data already exist [e.g., the package 'pavo' for R; @maia_pavo_2013]. While no program can compensate for having a clear understanding of the analysis at hand, programming languages are flexible by nature. This allows individual researchers to rapidly explore new methods as they are published, and methods developers to implement their analyses to increase the reach and impact of their work. A further advantage of programmatic analyses is that analysis scripts, when properly curated, represent a complete documented history of a study's methods. The chain of modification of raw data may thus be preserved, both for researchers revisiting their own work in the future, and for other scientists looking to build directly on the results of previous studies.

@@

Our review of the recent literature highlights some impediments to reproducibility in the study of biological colourationcoloration. The simplest step towards reproducible research is through the comprehensive reporting of methods, yet key information is often unreported (Fig. 1). Here we provide a list of methodological information that we suggest be specified in studies that focus on the spectrometric- or camera-based analysis of colour (Tables 1 & 2). In addition to the suggested information, we also recommend that raw spectra be presented where possible, as these figures allow for the rapid assessment of the nature and quality of the data independent of downstream processing. We also emphasise the importance of explicitly outlining the biological justification underlying all choices for data capture and analysis (@kemp_integrative_2015). Finally, the public storage of data along with detailed meta-data plays a pivotal role in reproducibility, yet public data storage in studies of colour traits is not yet the norm. We therefore encourage researchers to consider the substantial benefits of publicly archiving their data [@whitlock_data_2011; @piwowar_sharing_2007]. Overall, we hope that our guidelines will encourage researchers to think about the reproducibility of their work and the advantages of increased transparency, which will continue the advance of an exciting era in colour research.
